---
title: "Test BayesPrism using a sample Seurat Object"
html:
    embed-resources: true
editor: visual
---

# AIM

test the tool with a real dataset

# Load the libraries

```{r}
#| warning: false
#| message: false

# 
library(Seurat)
library(presto)
library(BayesPrism)
library(tidyverse)
```

# Read in the data

Read in the reference sc dataset

```{r}
seurat_ref <- readRDS("/mnt/SSD01/training/routines/scRNAseq/R/Seurat/out/object/100_ifnb_DonorStim.rds")

# explore the metadata
seurat_ref@meta.data %>%
  head()

# plot the dataset with the relative annotaiton
DimPlot(seurat_ref,group.by = "seurat_annotations")
```

Read in the sample pBulk generated. This is just one sample generated by summing up all the cells form the reference dataset. I have noticed that to avoid errors I need to have need at least two samples in the bulk dataset. To avoid any issue, and keep the example simple, I have replicated the same sample two time.

```{r}
# bulk_counts <- readRDS("../out/object/rawCount_pBulk_treat_100_ifnb.rds")
bulk_counts <- readRDS("../../out/object/rawCount_pBulk_all_100_ifnb.rds") %>%
  cbind(readRDS("../../out/object/rawCount_pBulk_all_100_ifnb.rds"))

colnames(bulk_counts) <- c("V1","V2")

head(bulk_counts)
```

Now produce the expected fraction per cell type for the bulk dataset.

```{r}
# LUT_prop <- seurat_ref@meta.data %>%
#   group_by(seurat_annotations,stim) %>%
#   summarise(n = n(),.groups = "drop") %>%
#   group_by(stim) %>%
#   mutate(tot = sum(n),
#          prop = n/tot)

LUT_prop <- seurat_ref@meta.data %>%
  group_by(seurat_annotations) %>%
  summarise(n = n(),.groups = "drop") %>%
  # group_by(stim) %>%
  mutate(tot = sum(n),
         prop = n/tot)

LUT_prop
```

Check the dimensions of the data.

```{r}
dim(seurat_ref)
dim(bulk_counts)
```

# Prepare Data for BayesPrism

BayesPrism requires the data in specific formats. We need to extract the raw counts and cell labels from our Seurat object and ensure the gene names match between our reference and bulk data (same annotation).

```{r}
# Extract Information from the Seurat Object
# BayesPrism works best with raw counts.
# we can input sparse matrices: https://github.com/Danko-Lab/BayesPrism/issues/58
sc_counts_matrix <- GetAssayData(seurat_ref, slot = "counts")

# Get cell type labels. These will be our 'cell.type.labels'.
cell_types <- seurat_ref$seurat_annotations

# Get cell state labels. For a simple case, these can be the same as cell types.
# In more complex analyses, these could be cluster IDs or subtypes.
# https://github.com/Danko-Lab/BayesPrism/issues/66
cell_states <- seurat_ref$seurat_annotations

# reshape the data to be accepted by the tool
# generally data are reporeted as feature X sample, but in the tool they are handled as sample X feature.
sc.dat <- t(sc_counts_matrix)
sc.dat[1:5,1:5]
dim(sc.dat)
class(sc.dat)

bk.dat <- t(bulk_counts)
bk.dat[,1:5]
dim(bk.dat)
class(bk.dat)

```

# Filter outlier genes from scRNA-seq data

Now it is possible to run the reccommended filtering on the genes. This step can be skipped, but it might affect the final quality of the dataset. We remove the genes from selected groups. Note that when sex is not identical between the reference and mixture, we recommend excluding genes from chrX and chrY. We also remove lowly transcribed genes, as the measurement of transcription of these genes tend to be noise-prone. Removal of these genes can also speed up computation.

```{r}
#
sc.dat.filtered <- cleanup.genes(input = sc.dat,
                                 input.type = "count.matrix",
                                 species = "hs",
                                 gene.group = c( "Rb","Mrp","other_Rb","chrM","MALAT1","chrX","chrY") ,
                                 exp.cells = 5)

# compare the dimensions before and after filtering
dim(sc.dat.filtered)
dim(sc.dat)
```

Next, we check the concordance of gene expression for different types of genes. As bulk and single cell data are usually collected by different experimental protocols, they may have different sensitivity to different types of genes. Note this function only works for human data. For other species, you are advised to make plots by yourself.

```{r}
#| eval: false

plot.bulk.vs.sc(sc.input = sc.dat.filtered,
                bulk.input = bk.dat
                #pdf.prefix="gbm.bk.vs.sc" specify pdf.prefix if need to output to pdf
)
```

Generally we observe that protein coding genes are the most concordant group between assays. To reduce batch effects and speed up computation, we perform deconvolution on protein coding genes. We have also tried to run BayesPrism on all genes. The results were similar. Subset protein coding genes.

```{r}
#
sc.dat.filtered.pc <-  select.gene.type(sc.dat.filtered,
                                        gene.type = "protein_coding")

# check the dimensions
str(sc.dat.filtered.pc)
dim(sc.dat.filtered.pc)
dim(sc.dat.filtered)
dim(sc.dat)
```

Optionally, at this point, it is possible to run the DE to select markers gene per cell_type. The idea is to further filter the reference dataset to focus only in highly specific genes. In this example I am not running this step.

```{r}
#| eval: false

# set the Idents for the cell types
Idents(seurat_ref) <- "seurat_annotations"

# find the markers genes using the default parameters
diff.exp.stat <- FindAllMarkers(seurat_ref)

str(diff.exp.stat)

# Ideally, we would like to have sufficient number of genes selected for each cell type (>50). If not, users may lower the cutoff. To subset our count matrix over the signature genes we can proceed as follows

marker_genes <- diff.exp.stat %>%
  # apply any filter here
  # mutate(delta.pct = pct.1 - pct.2) %>%
  # filter(avg_log2FC > 0.1,p_val_adj < 0.01,abs(delta.pct)>0.1) %>%
  pull(gene) %>%
  unique()

# match the marker genes with the available genes
filter_genes <- intersect(marker_genes,colnames(sc.dat.filtered.pc))

# how many genes are retained
length(filter_genes)

# filter the matrix
sc.dat.filtered.pc.sig <- sc.dat.filtered.pc[,filter_genes]

dim(sc.dat.filtered.pc.sig)
sc.dat.filtered.pc.sig[1:10,1:10]
```

# Run BayesPrism Deconvolution

With the data properly formatted and aligned, we can now run the deconvolution. Construct the BayesPrism Object by combineing the reference and cell labels into a 'prism' object. BayesPrism works for both tumor and non-tumor datasets. Many of the benchmarks in the manuscript were done on non-tumor datasets, such PBMC and brain cells. Simply set `key=NULL` when constructing the prism object.

```{r}
my_prism <- new.prism(
  reference = sc.dat.filtered.pc,
  mixture = bk.dat,
  # No specific tumor/normal key needed for this example: https://github.com/Danko-Lab/BayesPrism/issues/76
  key = NULL,
  cell.type.labels = cell_types,
  cell.state.labels = cell_states,
  input.type = "count.matrix"
)
```

Run the Deconvolution Algorithm. This is the core computational step. We'll use 20 cores for this example. Adjust `n.cores` as needed.

```{r}
#| eval: false

# for quarto file, skip the processing and load the result
bp_results <- run.prism(prism = my_prism, n.cores = 20)
```
```{r}
#| include: false

bp_results <- readRDS("../../out/object/bp_results_rawCount_pBulk_all_100_ifnb.rds")
```

You can look at the returned object to see what it contains

```{r}
bp_results

# saveRDS(bp_results,"../out/object/bp_results_rawCount_pBulk_treat_100_ifnb.rds")
# saveRDS(bp_results,"../../out/object/bp_results_rawCount_pBulk_all_100_ifnb.rds")
```

# Extract and Calculate Deconvoluted Reads and fractions

Deconvoluted reads can be extracted as follow. Notice that they are not integer, but according to the authors, it is possible to round them to integers (https://github.com/Danko-Lab/BayesPrism/issues/121).

```{r}
# x <- "Endothelial"

# deconvoluted estimates
list_bulk <- lapply(unique(cell_types), function(x){
  mat <- get.exp(bp=bp_results,
                 state.or.type="type",
                 cell.name=x)
  
  data <- mat %>%
    t()
  
  return(data)
}) %>%
  setNames(unique(cell_types))

str(list_bulk)

# generate integer estimates
list_bulk_integer <- lapply(unique(cell_types), function(x){
  mat <- get.exp(bp=bp_results,
                 state.or.type="type",
                 cell.name=x)
  
  # generate integer level estimates
  # https://github.com/Danko-Lab/BayesPrism/issues/121
  # round up the Z matrix to the nearest integer and use it as input for DESeq2 (without any normalization step).
  data <- mat %>%
    t() %>%
    round()
  return(data)
}) %>%
  setNames(unique(cell_types))

str(list_bulk_integer)

```

Notice that the number of genes matches is smaller than the input.

```{r}
lapply(list_bulk, function(x){
  dim(x)
})

dim(sc.dat.filtered.pc)
```

Get the fraction estimated

```{r}
theta <- get.fraction(bp=bp_results,
                      which.theta="final",
                      state.or.type="type")
theta
```

# Test

Compare the expected prop vs the actual ones

```{r}
# df_cor <- full_join(
#   LUT_prop,
#   theta %>%
#     as.data.frame() %>%
#     rownames_to_column("sample_id") %>%
#     pivot_longer(names_to = "cell_type",values_to = "prop",-sample_id),
#   by = c("stim" = "sample_id","seurat_annotations"="cell_type"),suffix = c(".expexted",".BayesPrism"))

df_cor <- full_join(
  LUT_prop,
  theta %>%
    as.data.frame() %>%
    rownames_to_column("sample_id") %>%
    pivot_longer(names_to = "cell_type",values_to = "prop",-sample_id),
  by = c("seurat_annotations"="cell_type"),suffix = c(".expexted",".BayesPrism"))

df_cor %>%
  ggplot(aes(x=prop.expexted,y=prop.BayesPrism)) +
  geom_point() +
  geom_abline(intercept = 0,slope = 1,col="red",linetype = "dashed") +
  theme_bw() +
  # facet_wrap(~seurat_annotations)+
  theme(strip.background = element_blank())

```

Confirm that the sum of the deconvoluted pBulk per sample correponds to the actual input values

```{r}
# sum all the deconvoluted matrices obtained
bulk_convoluted <- purrr::reduce(list_bulk,`+`)

str(bulk_convoluted)

# the two matrices are not the same as the convoluted version is missing the genes that have been filtered out
dim(bulk_counts)
dim(bulk_convoluted)

# match the gene id in both objects before making the comparison
gene_id <- rownames(bulk_convoluted)
str(bulk_convoluted)
str(bulk_counts[gene_id,] %>% as.matrix())

# they are the same
all.equal(bulk_convoluted,bulk_counts[gene_id,] %>% as.matrix())

```
